Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars

### 2019-11-30 project/overleaf/paper.tex
Paper v3.5.0
Version sent to MNRAS

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[firstpage]{draftwatermark}
\SetWatermarkText{DRAFT}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{The Sun lithium problem and how magnetic braking could help to solve it}

@@

Theoretical models of stars do not inform about the initial amount of Li that a star has, they only describe how it is exhausted. Therefore, in order to make an accurate estimate of the initial abundance of Li, it is a prerequisite to be able to compare observations and models beforehand. Our Sun represents a unique exception, since it allows us to know the current abundance of this element in its photosphere, $A(\isotope[7]{Li}) = 1.1 \pm 0.1 \, ex$ \citep{Jeffries2004}, where $A(\isotope[7]{Li})$ is defined according to Eq.~\ref{eq:A_Li}\par

@@

Making using of ($\eta_*$), the $R_{A}$ is defined as the point in which the magnetic field energy density $B^{2}/8\upi$, marked by magnetic field strength ($B$), and the kinetic energy density $\frac{1}{2}\rho\nu^{2}$, defined by the mass density ($\rho$) and velocity of the stellar wind $(\nu)$, are balanced. That is, where the coefficient between both is equal to 1. In case that $R_{A}$ is greater than the star radius, then the wind flow will have to follow the magnetic field. As a consequence, the material will leave the stellar surface with a higher specific AM, as the co-rotation radius has increased and it will roughly correspond to the $R_{A}$. This effect is known as MB.\par

Each of these factors has a decisive influence on the future of the star and therefore in our routine implementation. On the one hand, stars with similar initial masses but different mass loss ($\Dot{M}$) ratios will end up evolving very differently. On the other hand, the ionized particles carried by the solar wind not only contribute to the loss of mass but also to the loss of kinetic ($K_e$) energy that is deposited in the interstellar medium. Given a star with a spherically symmetric wind, $\Dot{M}$ and $K_e$ are characterized and related by the following expressions:

@@

In turn, of the observable characteristics of the solar wind, we especially care about the parameters related to the amount of mass that is able to pull out of the outer layers of the star in a given time interval, and the speed that reaches the wind itself measured at a great distance from the star, the terminal velocity ($\nu_\infty$). In general, a magnetically channeled outflow will have a complex stream geometry but for convenience, the Eq.~\ref{eq:mass_loss} simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law, where $r$ represents the distance from the center of the star to the point where the velocity is to be measured, and $R_*$ is the radius of the star:
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-R_*/r)
\end{equation}
\end{ceqn}
where $\nu_\infty$ is the terminal wind velocity defined as the velocity that the wind, the outflowing matter, reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

@@

Given that the version of MESA used in this work does include neither the treatment of surface magnetic fields (it does account for transport by internal magnetic fields of AM and chemical elements) and therefore, nor for magnetic braking effects, it is necessary to extend the simulator to include this phenomenon. In the present work we have followed the theoretical work on MB in the Sun carried out by \citet{Weber1967}, making use of the expression \ref{eq:j_dot} for the calculation of the AML.
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A} \label{eq:j_dot}
\end{equation}
\end{ceqn}
where $\Dot{M}$ is the mass loss rate, $\Omega$ the angular velocity at star surface and $R_A$ the Alfv\'{e}n radius. \par

The above expression can be rewritten so that it depends on the $\eta_*$ instead of the $R_A$ as follows:
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{*}\eta_* \label{eq:j_dot_mesa}
\end{equation}
\end{ceqn}
which is a more convenient expression to be implemented in MESA because is based on values directly exposed during the simulations and convenient for the approach described in \citet{Ud-Doula2007}.

@@

In its simplest form the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial rotation rate ($\Omega$) at star surface. The numerical simulations trace the rotational history and A(Li) of a $1\, \msun$ star for a variety of initial values for $B$ and $\Omega$. Although stars have a three-dimensional structure and the equations that model their evolution must take this fact into account,it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two \citep[shellular approximation,][]{Meynet1997}. MESA uses this simplification to accelerate computational calculations \citet{Paxton2013}.\par

@@

We define here also additional conventions and assumptions adopted in the simulations. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, \mathrm{pr} = 0.0142$) not the current photospheric metallicity ($Z = Z_{\odot, \mathrm{ph} = 0.0134$) as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $\rsun = 6.957x10^{10}\, cm$ and $\msun = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. That work, in particular its section 4.1, has been used as a starting point for ours regarding the parameterization of the MESA project which is calibrated to reproduce the measured element abundances on the solar surface.\par

@@

The enumerated effects \ref{itm:1} and \ref{itm:2} are standard features offered by MESA. On the contrary, the effect \ref{itm:3} is not available as an out-of-the-box feature during the elaboration of this paper and represents our extension to MESA. Additionally, we do not take into account either the influence of internal magnetic fields or their existence during the T-Tauri phase. We adopt a simple and pragmatic theoretical approach to establish their occurrence when the star is reaching the ZAMS phase. The adopted criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine will be activated acting as an additional mechanism to those existing in the MESA evolutionary code that participates in the star AML. Additionally, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches the TAMS.\par

In this study we do not only focused in investigating the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars. Keeping in mind this approach, our major objective is aimed to check the degree of validity of the resulting model after including the MB effect as a complementary mechanism of AML that can contribute to explain the evolution of Li in the Sun and other solar-type stars. A set of different simulation scenarios with magnetic field strengths ranging from between 3.0 and 5.0 G have been exercised. We'd like to stress that the magnetic field remains constant during the PMS \& MS phases. We compute the evolution of $1\msun$ stellar models at solar metallicity with $\oomegac$ varying between $0.0084$ and $0.0336$. More advanced models for more complex arrangements of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

@@

Figure~\ref{fig:li_var_vel_0g} shows the evolution of surface Li abundance relative to H as a function of time and for several $1\msun$ models initialized with different rotational velocities. The simulations take into account the effects of rotation and AML caused by stellar winds but not those of MB. The purple star and square are surface Li abundances for the present-day Sun \citep{Asplund2009} and Pleiades \citep{Sestito2005}.\par

Additionally, notice how the Li abundance in the star surface decreases over time for all simulated models. The solid black line represents the reference model that adopts the solar-calibrated envelope overshooting parameters as documented in \citet{Choi2016}. All models burn too much Li before the ZAMS and therefore do not match with the Pleiades average surface Li abundance. Also remarkable is the fact that there are barely any differences between the different models in terms of the abundance of Li for much of the PMS. Only after one million years is Li destroyed to an accentuated degree, since the necessary temperature is not reached at BCZ before. Afterwards the different models destroy the Li in a very similar way due to two main reasons. On the one hand, the convective zones that the models develop have very similar sizes, so that the temperature at the BCZ is practically the same. On the other hand, the differential rotation between the core and the convective zone does not begin to develop until it reaches $ \approx 10^6$ years, achieving its maximum difference in the ZAMS about $ \approx 10^7$ years. It is at this moment that the combined effects of turbulence and rotational difference should increase, leading to notable differences in the evolution of A(Li) (see Figure~\ref{fig:li_var_vel_0g_z1}). Later, the reference model does not deplete Li efficiently on the MS and fail (again) to match the current solar surface Li abundance. On the other hand, the other models which include rotation during the PMS with values of $\oomegac$ between 0.0084 and 0.0336 are able to burn Li in a more realistic way although only one of them (green line) is close to match the present-day Li abundance of the Sun but its rotational velocity is much higher (see Figure~\ref{fig:rot_vel_0g}) than the $2\,\kms$ of the Sun \citep{Gill2012}. \par

For much of the PMS the star rotates as a solid body (see Figure~\ref{fig:rot_vel_0g}) and this is because the star has a completely convective interior. It is not until the end of Hayashi track that the star begins to develop a radiative core due to the increase in temperature that occurs inside the star during the contraction process. It is in this stage of evolution of the star when a difference in angular velocity appears between the upper and lower limits of the radiative and convective zones respectively. The degree of differential rotation is directly influenced by the initial $\Omega$. As the models are initialized with a higher angular velocity, the difference in velocity between the BCZ and the star surface is accentuated in a directly proportional mode; the higher the initial velocity, the bigger the velocity gradient between the bottom and top limits of the CZ. As a consequence, the turbulence strength located at the BCZ increases so that Li can reach regions with temperatures about $\tli$ where it is finally burned and destroyed (see figure~ \ref{fig:li_var_vel_0g}). Other investigations \citep{Bouvier2018, Baraffe2017}  point to a tendency diametrically opposed to the one exposed here, i.e. the faster the rotation speed, the greater the abundance of Li on the surface of the star. \par

\begin{figure}
	\includegraphics[trim = 35mm 15mm 20mm 15mm, clip,width=\columnwidth]{figures/li_var_vel_0_0g_z1.eps}
    \caption {Similar to Figure~\ref{fig:li_var_vel_0g} but zooming in on the ZAMS. The models with a higher initial rotational velocity reach already the ZAMS with a lower amount of Li measure on the star surface.}
    \label{fig:li_var_vel_0g_z1}
\end{figure}

\begin{figure}
	\includegraphics[trim = 30mm 15mm 20mm 15mm, clip,width=\columnwidth]{figures/rot_vel_var_vel_0_0g.eps}
    \caption{The evolution of angular velocity at surface (solid line) and the top (dashed line) and bottom (dotted line) limits of the uppermost convective zone, as a function of time for several 1 $\msun$ models. The models includes PMS rotation with $\oomegac$ values between 0.0084 and 0.0336. The purple star is the surface angular velocity for the present-day Sun \citep{Gill2012}. The dashed vertical lines make reference to the ZAMS.}
    \label{fig:rot_vel_0g}
\end{figure}

Another well known structural effect of rotation is the decrease of the effective temperature ($\teff$) and to a less extent of stellar luminosity ($L$). Both effects can be observed graphically in the HR diagram of Figure~\ref{fig:hr_var_vel_0g} which shows a zoomed-in view of evolutionary tracks from the ZAMS until the TAMS for several $1\msun$ models initialized with different rotational velocities. If we compare the non-rotating model (black solid line on the left side) with the rotating ones we can recognize that at the end of the PMS, the latter reach before the ZAMS and exhibit a lower $\teff$ than the former. These results are in line with those of previous studies \citep[see e.g. ][]{Eggenberger2012,Piau2001,Pinsonneault1989}.\par
%(see e.g. \citet{Eggenberger2012}, \citet{Piau2001}, \citet{Pinsonneault1989}).\par

@@

Figure \ref{fig:li_var_vel_4_0g} shows the evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H}, as a function of time for several 1 $\msun$ models. Those models were initialized with different rotational velocities and took into consideration both the effects of AML and the MB caused by a magnetic field of intensity 4G. If we compare it with Figure~\ref{fig:li_var_vel_0g} in which the effects of MB are neglected, we notice how the profiles of Li abundance are altered during PMS and MS. In the first phase we can describe the effect as modest, somewhat expected and in line with the fact that the AML caused by MB (see Eq.~\ref{eq:j_dot}) depends directly on the mass loss rate. If we take into account that for solar-type stars the models predict a modest total mass loss rate, that value is even much lower during PMS. On the contrary, during the MS it is observed that the AML is much more significant, causing the star to rotate more slowly and as a result a smaller amount of Li to be destroyed.\par

If we concentrate our attention on Figure~\ref{fig:li_var_vel_4_0g} which represents the evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H}, as a function of time for several 1 $\msun$ models initialized with different rotational velocities and considers the effects of AML and the MB caused by a magnetic field of intensity 4G, and compare it with Figure~\ref{fig:li_var_vel_0g} in which the effects of MB are neglected, we will see how the profiles of Li abundance are altered during PMS and MS. In the first phase we can describe the effect as modest, somewhat expected and in line with the fact that the AML caused by MB (see Eq.~\ref{eq:j_dot}) depends directly on the mass loss rate. If we take into account that for solar-type stars the models predict a modest total mass loss rate, that value is even much lower during PMS. On the contrary, during the MS it is observed that the AML is much more significant, causing the star to rotate more slowly and as a result a smaller amount of Li to be destroyed.\par

@@

The effect of the MB routine can be seen even more clearly in figures \ref{fig:rot_vel_4g}, \ref{fig:rot_vel_4g_z1} \& \ref{fig:grid_rot_vel}. In those figures we represent the rotation profiles for the surface of the stars and for the limits of the convective envelope for a number of 1 $\msun$ models initialized with different rotational velocities and considering the influence of MB. Similarly to the evolution profiles of the Li on the surface commented in the previous paragraph, the effect of the routine is visible once the ZAMS is reached. If we compare the evolution of the curves presented here with those of Figure~\ref{fig:rot_vel_0g} we see how the star, instead of continuing to increase its $\Omega$, begins to slow down after having reached its maximum in the passage through the ZAMS. It is also at this stage when the angular velocities at the surface of the star and at the BCZ reach their maximum difference. On the other hand, the effect of magnetic braking causes the angular velocities between both zones of the star to decrease until, for an age close to that of the Sun (see Figure~\ref{fig:rot_vel_4g_z1}), the star practically rotates like a rigid solid. These results are also consistent with those obtained by \citet{Eggenberger2010} as far as the effect of the magnetic field, in particular its influence on the loss of angular momentum, has on the rotational velocity of the star. In our models we have also chosen to include the rotation effects offered by MESA but not those of Spruit-Tayler (ST) dynamo on the diffusion of elements. The reason is that the implementation of MESA that makes of the ST dynamo is to simulate internal magnetic fields and not for surface magnetic fields \citep{Paxton2013}.

In a similar way we also observe that the models with lower angular rotation generally end up exhibiting higher values in the abundance of Li on the surface (see figures \ref{fig:li_var_vel_4_0g}, \ref{fig:grid_li_var_vel} \& \ref{fig:grid_li_var_g}). In none of those cases we do obtain values of Li on the surface higher than those shown by the model without rotation.\par

@@

\begin{figure}
	\includegraphics[trim = 30mm 15mm 20mm 15mm, clip,width=\columnwidth]{figures/rot_vel_var_vel_4_0g_z1.eps}
    \caption{Similar to Figure~\ref{fig:rot_vel_4g} but now showing in detail the surface rotational velocity as the star approaches TAMS.}
    \label{fig:rot_vel_4g_z1}
\end{figure}

It is also remarkable to observe that, for some of the models depicted in Figure~\ref{fig:rot_vel_4g}, their $\Omega$ at the star surface (solid line) is slower than the one of the upper limit of their convective layer (dashed line). This fact is most probably a consequence of the expansion of the superficial layers after the mass loss produced by stellar winds.\par

The MB also leaves its mark on the HR diagram by significantly affecting the $\teff$ of the star. To visualize this effect we will take as reference Figure~\ref{hr_vc_0336_var_g_z1} in which all models were started with the same value $\oomegac=0.0336$ but the intensity of the simulated magnetic field was different. In particular, let us pay attention to the evolutionary trace in which a magnetic field was not simulated (black solid line) and to the rest of the evolutionary traces in which it was. We have that the group of simulations with presence of a magnetic field produces hotter stars due to the MB influence. The lower speed with which the star rotates due to the MB effect causes the increase of the $\teff$, being this difference of practically $95\,\Kelvin$ between the simulated models with $0\,\Gauss$ and $5.0\,\Gauss$ respectively for $log(\llsun)=0$.

\begin{figure}
	\includegraphics[trim = 30mm 15mm 15mm 15mm, clip,width=\columnwidth]{figures/hr_vc_0336_var_g_z1.eps}
    \caption{Similar to Figure~\ref{fig:hr_var_vel_0g} but now showing in detail the effects of magnetic braking on the evolutionary tracks for different magnetic field strengths and $\oomegac=0.0336$. The presence of a magnetic field produces hotter stars due to the influence of the magnetic braking on the rotational velocity of the star.}
    \label{hr_vc_0336_var_g_z1}
\end{figure}

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
%\usepackage[firstpage]{draftwatermark}
%\SetWatermarkText{DRAFT}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars}

@@

Theoretical models of stars do not inform about the initial amount of Li that a star has, they only describe how it is exhausted. Therefore, in order to make an accurate estimate of the initial abundance of Li, it is a prerequisite to be able to compare observations and models beforehand. Our Sun represents a unique exception, since it allows us to know the current abundance of this element in its photosphere, $A(\isotope[7]{Li}) = 1.1 \pm 0.1 \, dex$ \citep{Jeffries2004}, where $A(\isotope[7]{Li})$ is defined according to Eq.~\ref{eq:A_Li}\par

@@

Making use of $\eta_*$, the $R_{A}$ is defined as the point in which the magnetic field energy density $B^{2}/8\upi$, marked by magnetic field strength ($B$), and the kinetic energy density $\frac{1}{2}\rho\nu^{2}$, defined by the mass density ($\rho$) and velocity of the stellar wind $(\nu)$, are balanced. That is, where the coefficient between both is equal to 1. In case that $R_{A}$ is greater than the star radius, then the wind flow will have to follow the magnetic field. As a consequence, the material will leave the stellar surface with a higher specific AM, as the co-rotation radius has increased and it will roughly correspond to the $R_{A}$. This effect is known as MB.\par

Each of these factors has a decisive influence on the future of the star and therefore in our routine implementation. On the one hand, stars with similar initial masses but different mass loss ($\Dot{M}$) ratios will end up evolving very differently. On the other hand, the ionized particles carried by the solar wind not only contribute to the mass loss but also to the loss of kinetic ($K_e$) energy that is deposited in the interstellar medium. Given a star with a spherically symmetric wind, $\Dot{M}$ and $K_e$ are characterized and related by the following expressions:

@@

In turn, of the observable characteristics of the solar wind, we especially care about the parameters related to the amount of mass that is able to pull out of the outer layers of the star in a given time interval, and the speed that reaches the wind itself measured at a great distance from the star, the terminal velocity ($\nu_\infty$). In general, a magnetically channeled outflow will have a complex stream geometry but for convenience, the Eq.~\ref{eq:mass_loss} simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law (Eq.~\ref{eq:vel_law}), where $r$ represents the distance from the center of the star to the point where the velocity is to be measured, and $R_*$ is the radius of the star:
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-R_*/r) \label{eq:vel_law}
\end{equation}
\end{ceqn}
where $\nu_\infty$ is the terminal wind velocity defined as the velocity that the wind, the outflowing matter, reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

@@

Given that the version of MESA used in this work does include neither the treatment of surface magnetic fields (it does account for transport by internal magnetic fields of AM and chemical elements) and therefore, nor for magnetic braking effects, it is necessary to extend the simulator to include this phenomenon. In the present work we have followed the theoretical work on MB in the Sun carried out by \citet{Weber1967}, making use of Eq.~ \ref{eq:j_dot} for the calculation of the AML.
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A} \label{eq:j_dot}
\end{equation}
\end{ceqn}
where $\Dot{M}$ is the mass loss rate, $\Omega$ the angular velocity at star surface and $R_A$ the Alfv\'{e}n radius. \par

The above expression can be rewritten so that it depends on the $\eta_*$ instead of the $R_A$ as follows:
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{*}\eta_* \label{eq:j_dot_mesa}
\end{equation}
\end{ceqn}
which is a more convenient expression to be implemented in MESA because is based on values directly exposed during the simulations and suitable for the approach described in \citet{Ud-Doula2007}.

@@

In its simplest form the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial rotation rate ($\Omega$) at star surface. The numerical simulations trace the rotational history and A(Li) of a $1\, \msun$ star for a variety of initial values for $B$ and $\Omega$. Although stars have a three-dimensional structure and the equations that model their evolution must take this fact into account, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two \citep[shellular approximation,][]{Meynet1997}. MESA uses this simplification to accelerate computational calculations \citep{Paxton2013}.\par

@@

We define here also additional conventions and assumptions adopted in the simulations. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, \mathrm{pr} = 0.0142}$) not the current photospheric metallicity ($Z = Z_{\odot, \mathrm{ph} = 0.0134}$) as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $\rsun = 6.957x10^{10}\, cm$ and $\msun = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. That work, in particular its section 4.1, has been used as a starting point for ours regarding the parameterization of the MESA project which is calibrated to reproduce the measured element abundances on the solar surface.\par

@@

The enumerated effects \ref{itm:1} and \ref{itm:2} are standard features offered by MESA. On the contrary, the effect \ref{itm:3} is not available as an out-of-the-box feature during the elaboration of this paper and represents our extension to MESA. Additionally, we do not take into account either the influence of internal magnetic fields or their existence during the T-Tauri phase. We adopt a simple and pragmatic theoretical approach to establish their occurrence when the star is reaching the ZAMS phase. The adopted criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine will be activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participates in the star AML. Additionally, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches the TAMS.\par

In this study we do not only focus on investigating the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars. Keeping in mind this approach, our major objective is aimed to check the degree of validity of the resulting model after including the MB effect as a complementary mechanism of AML that can contribute to explain the evolution of Li in the Sun and other solar-type stars. A set of different simulation scenarios with magnetic field strengths ranging between 3.0 and 5.0 G have been exercised. We would like to stress that the magnetic field remains constant during the PMS \& MS phases. We compute the evolution of $1\msun$ stellar models at solar metallicity with $\oomegac$ varying between $0.0084$ and $0.0336$. More advanced models for more complex arrangements of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

@@

Figure \ref{fig:li_var_vel_0g} shows the evolution of surface Li abundance relative to H as a function of time and for several $1\msun$ models initialized with different rotational velocities. The simulations take into account the effects of rotation and AML caused by stellar winds but not those of MB. The purple star and square are the surface Li abundances for the present-day Sun \citep{Asplund2009} and for the Pleiades, respectively.  \citep{Sestito2005}.\par

Additionally, notice how the Li abundance in the star surface decreases over time for all simulated models. The solid black line represents the reference model that adopts the solar-calibrated envelope overshooting parameters as documented in \citet{Choi2016}. All models burn too much Li before the ZAMS and therefore do not match with the Pleiades average surface Li abundance. Also remarkable is the fact that there are barely any differences between the different models in terms of the abundance of Li for much of the PMS. Only after one million years is Li destroyed to an accentuated degree, since the necessary temperature is not reached at BCZ before. Afterwards the different models destroy the Li in a very similar way due to two main reasons. On the one hand, the convective zones that the models develop have very similar sizes, so that the temperature at the BCZ is practically the same. On the other hand, the differential rotation between the core and the convective zone does not begin to develop until it reaches $ \approx 10^6$ years, achieving its maximum difference in the ZAMS about $ \approx 10^7$ years. It is at this moment that the combined effects of turbulence and rotational difference should increase, leading to notable differences in the evolution of A(Li) (see Figure \ref{fig:li_var_vel_0g_z1}). Later, the reference model does not deplete Li efficiently on the MS and fail (again) to match the current solar surface Li abundance. On the other hand, the other models which include rotation during the PMS with values of $\oomegac$ between 0.0084 and 0.0336 are able to burn Li in a more realistic way although only one of them (green line) is close to match the present-day Li abundance of the Sun but its rotational velocity is much higher (see Figure \ref{fig:rot_vel_0g}) than the $2\,\kms$ of the Sun \citep{Gill2012}. \par

For much of the PMS the star rotates as a solid body (see Figure \ref{fig:rot_vel_0g}) and this is because the star has a completely convective interior. It is not until the end of Hayashi track that the star begins to develop a radiative core due to the increase in temperature that occurs inside the star during the contraction process. It is in this stage of evolution of the star when a difference in angular velocity appears between the upper and lower limits of the radiative and convective zones respectively. The degree of differential rotation is directly influenced by the initial $\Omega$. As the models are initialized with a higher angular velocity, the difference in velocity between the BCZ and the star surface is accentuated in a directly proportional mode; the higher the initial velocity, the bigger the velocity gradient between the bottom and top limits of the CZ. As a consequence, the turbulence strength located at the BCZ increases so that Li can reach regions with temperatures about $\tli$ where it is finally burned and destroyed (see Figure  \ref{fig:li_var_vel_0g}). Other investigations \citep{Bouvier2018, Baraffe2017}  point to a tendency diametrically opposed to the one exposed here, i.e. the faster the rotation speed, the greater the abundance of Li on the surface of the star. \par

\begin{figure}
	\includegraphics[trim = 35mm 15mm 20mm 15mm, clip,width=\columnwidth]{figures/li_var_vel_0_0g_z1.eps}
    \caption {Similar to Figure \ref{fig:li_var_vel_0g} but zooming in on the ZAMS. The models with a higher initial rotational velocity reach already the ZAMS with a lower amount of Li measure on the star surface.}
    \label{fig:li_var_vel_0g_z1}
\end{figure}

\begin{figure}
	\includegraphics[trim = 30mm 15mm 20mm 15mm, clip,width=\columnwidth]{figures/rot_vel_var_vel_0_0g.eps}
    \caption{The evolution of angular velocity at surface (solid line) and at the bottom (dotted line) limit of the uppermost convective zone, as a function of time for several 1 $\msun$ models. The models includes PMS rotation with $\oomegac$ values between 0.0084 and 0.0336. The purple star is the surface angular velocity for the present-day Sun \citep{Gill2012}. The dashed vertical lines make reference to the ZAMS.}
    \label{fig:rot_vel_0g}
\end{figure}

Another well known structural effect of rotation is the decrease of the effective temperature ($\teff$) and to a less extent of stellar luminosity ($L$). Both effects can be observed graphically in the HR diagram of Figure \ref{fig:hr_var_vel_0g} which shows a zoomed-in view of evolutionary tracks from the ZAMS until the TAMS for several $1\msun$ models initialized with different rotational velocities. If we compare the non-rotating model (black solid line on the left side) with the rotating ones we can recognize that at the end of the PMS, the latter reach before the ZAMS and exhibit a lower $\teff$ than the former. These results are in line with those of previous studies \citep[see e.g. ][]{Eggenberger2012,Piau2001,Pinsonneault1989}.\par
%(see e.g. \citet{Eggenberger2012}, \citet{Piau2001}, \citet{Pinsonneault1989}).\par

@@

Figure \ref{fig:li_var_vel_4_0g} shows the evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H}, as a function of time for several 1 $\msun$ models. Those models were initialized with different rotational velocities and took into consideration both the effects of AML and the MB caused by a magnetic field of intensity 4G. If we compare it with Figure \ref{fig:li_var_vel_0g} in which the effects of MB are neglected, we notice how the profiles of Li abundance are altered during PMS and MS. In the first phase we can describe the effect as modest, somewhat expected and in line with the fact that the AML caused by MB (see Eq.~\ref{eq:j_dot}) depends directly on the mass loss rate. If we take into account that for solar-type stars the models predict a modest total mass loss rate, that value is even much lower during PMS. On the contrary, during the MS it is observed that the AML is much more significant, causing the star to rotate more slowly and as a result a smaller amount of Li to be destroyed.\par

@@

The effect of the MB routine can be seen even more clearly in Figures~\ref{fig:rot_vel_4g}, \ref{fig:rot_vel_4g_z1} \& Appendixes \footnote{The appendices comprise a series of grids as a function of time and for several 1 $\msun$ models in which on the one hand, the evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H} for both variable magnetic field intensities and angular velocities and on the other hand, the evolution of surface rotational velocity are shown.}. In those figures we represent the rotation profiles for the surface of the stars and for the bottom of the convective envelope for a number of 1 $\msun$ models initialized with different rotational velocities and considering the influence of MB. Similarly to the evolution profiles of the Li on the surface commented in the previous paragraph, the effect of the routine is visible once the ZAMS is reached. If we compare the evolution of the curves presented here with those of Figure \ref{fig:rot_vel_0g} we see how the star, instead of continuing to increase its $\Omega$, begins to slow down after having reached its maximum in the passage through the ZAMS. It is also at this stage when the angular velocities at the surface of the star and at the BCZ reach their maximum difference. On the other hand, the effect of magnetic braking causes the angular velocities between both zones of the star to decrease until, for an age close to that of the Sun (see Figure \ref{fig:rot_vel_4g_z1}), the star practically rotates like a rigid solid. These results are also consistent with those obtained by \citet{Eggenberger2010} as far as the effect of the magnetic field, in particular its influence on the loss of angular momentum, has on the rotational velocity of the star. In our models we have also chosen to include the rotation effects offered by MESA but not those of Spruit-Tayler (ST) dynamo on the diffusion of elements. The reason is that the implementation of MESA that makes of the ST dynamo is to simulate internal magnetic fields and not for surface magnetic fields \citep{Paxton2013}.

In a similar way we also observe that the models with lower angular velocity generally end up exhibiting higher values in the abundance of Li on the surface (see Figures~ \ref{fig:li_var_vel_4_0g}, \ref{fig:grid_li_var_vel} \& \ref{fig:grid_li_var_g}). In none of those cases we do obtain values of Li on the surface higher than those shown by the model without rotation.\par

@@

\begin{figure}
	\includegraphics[trim = 30mm 15mm 20mm 15mm, clip,width=\columnwidth]{figures/rot_vel_var_vel_4_0g_z1.eps}
    \caption{Similar to Figure \ref{fig:rot_vel_4g} but now showing in detail the surface rotational velocity as the star approaches TAMS.}
    \label{fig:rot_vel_4g_z1}
\end{figure}

The MB also leaves its mark on the HR diagram by significantly affecting the $\teff$ of the star. To visualize this effect we will take as reference Figure \ref{hr_vc_0336_var_g_z1} in which all models were started with the same value $\oomegac=0.0336$ but the intensity of the simulated magnetic field was different. In particular, let us pay attention to the evolutionary trace in which a magnetic field was not simulated (black solid line) and to the rest of the evolutionary traces in which it was. We have that the group of simulations with presence of a magnetic field produces hotter stars due to the MB influence. The lower speed with which the star rotates due to the MB effect causes the increase of the $\teff$, being this difference of practically $95\,\Kelvin$ between the simulated models with $0.0\,\Gauss$ and $5.0\,\Gauss$ respectively for $log(\llsun)=0$.

\begin{figure}
	\includegraphics[trim = 30mm 15mm 15mm 15mm, clip,width=\columnwidth]{figures/hr_vc_0336_var_g_z1.eps}
    \caption{Similar to Figure \ref{fig:hr_var_vel_0g} but now showing in detail the effects of magnetic braking on the evolutionary tracks for different magnetic field strengths and $\oomegac=0.0336$. The presence of a magnetic field produces hotter stars due to the influence of the magnetic braking on the rotational velocity of the star.}
    \label{hr_vc_0336_var_g_z1}
\end{figure}

As described in section \ref{sec_mesh}, the MB routine distributes the total amount of AML calculated according to Eq.~\ref{eq:k_jdot} among the different layers that compose the CZ. In Figure \ref{fig:cz_vc_028_var_b} we can observe the evolution of the most external CZ normalized with respect to the radius of the star for several 1 $\msun$ models. The models were all initialized with $\oomegac=0.0336$ and magnetic field strengths varying between $0.0\,\Gauss$ and $5.0\,\Gauss$. In accordance with the established models of stellar evolution, in a solar-type star the CZ covers practically all of it for a large part of the PMS. As it approaches the ZAMS, the CZ is decreasing as a consequence of the appearance of a radiative core and maintains an approximately constant radius until the final stage of the MS. In this point it increases significantly as a response to the generalized expansion of the star's radius. Regarding the effect of MB on the size of the CZ, we observe that as the intensity of the magnetic field increases, the size of the CZ decreases (see Figure \ref{fig:cz_vc_028_var_b}). The radiative core pushes outward to include a rapidly increasing fraction of the stellar mass, making that the temperature at the CZ base drops below $\tli$. This effect is most evident during MS (see Figure \ref{fig:cz_vc_028_var_b_z1}). The decrease in the size of the CZ is in line with the fact that less Li is destroyed by causing less star material to reach areas with temperatures above $\tli$.\par
